---
title: "DonorSearch Insights"
output: html_notebook
---
Import libraries
```{r setup}
library(ggplot2)
library(gridExtra)
library(dplyr)
```

Data Prep:
Load data in select only columns that are needed. Then, reformat Wealth Based Capacity into a more readable format for plotting. Next, create a non-donor data frame containing only prospects. Lastly, create Constituency.Type and Donor.Type columns.
```{r data}
df <- read.csv(file = "../Data/DonorSearch Insights.csv")
constituents <- df[,c("Total.Gift.Amount", "Constituency.Type", "Pref_Year", "Estimated.Capacity", "Wealth.Based.Capacity", "Acquisition.Rank", "Retention.Rank", "Upgrade.Rank", "Lifetime.Value.Rank", "Acquisition.Score", "Retention.Score", "Upgrade.Score", "Lifetime.Value.Score")]
constituents$Wealth.Based.Capacity <- substr(constituents$Wealth.Based.Capacity, 5, nchar(constituents$Wealth.Based.Capacity))
constituents$Wealth.Based.Capacity <- gsub(" ", "\n", constituents$Wealth.Based.Capacity)
constituents[constituents$Wealth.Based.Capacity == "$10\nMillion\n-\n$50\nMillion", "Wealth.Based.Capacity"] = "$10 Million\n-\n$50 Million"
constituents[constituents$Wealth.Based.Capacity == "$50\nMillion\n-\n$100\nMillion", "Wealth.Based.Capacity"] = "$50 Million\n-\n$100 Million"
constituents[constituents$Wealth.Based.Capacity == "$100\nMillion+", "Wealth.Based.Capacity"] = "$100 Million+"
constituents$Wealth.Based.Capacity <- factor(constituents$Wealth.Based.Capacity, levels=c("Unable\nto\nRate", "$1\n-\n$2,499", "$2,500\n-\n$4,999", "$5,000\n-\n$9,999", "$10,000\n-\n$14,999", "$15,000\n-\n$24,999", "$25,000\n-\n$49,999", "$50,000\n-\n$99,999", "$100,000\n-\n$249,999", "$250,000\n-\n$499,999", "$500,000\n-\n$999,999", "$1,000,000\n-\n$4,999,999", "$5,000,000\n-\n$9,999,999", "$10 Million\n-\n$50 Million", "$50 Million\n-\n$100 Million", "$100 Million+"))
non_donors <- constituents[constituents$Acquisition.Rank != "",]
constituents <- constituents %>% mutate(Constituency.Type = if_else(Constituency.Type == "Alumni", "Alumni", "Non-Alumni"))
constituents <- constituents %>% mutate(Donor.Type = if_else(Total.Gift.Amount > 0, "Donor", "Non-Donor"))
```

Scatter plots:
Create individual scatter plots for each insight score by wealth capacity, colored by the corresponding insight rank.
```{r scatter}
acq_col <- c("#48C2CB", "#3394A8", "#1D6786", "#083963")
ret_col <- c("#F38531", "#D9632B", "#C04125", "#A61F1F")
upg_col <- c("#73BCDC", "#3C7C94", "#053C4B")
ltv_col <- c("#C1D8DF", "#94B4B9", "#679092")

score_scatter <- function(df, col_name, colors) {
  score_label <- paste(col_name, "Score", sep = " ")
  score_col <- gsub(" ", ".", score_label)
  rank_label <- paste(col_name, "Rank", sep = " ")
  rank_col <- gsub(" ", ".", rank_label)
  ggplot(df, aes_string(x = "Estimated.Capacity", y = score_col, color = rank_col)) + 
    geom_point() + 
    facet_grid(.~Wealth.Based.Capacity, scales="free") +
    scale_color_manual(values=colors) +
    labs(x = "Wealth Based Capacity", y = score_label, color = rank_label) + 
    theme_bw() + 
    theme(legend.position = "bottom",
          strip.text.x = element_text(size = 4), 
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          panel.grid.minor = element_blank(),
          panel.spacing.x = unit(0,"line"))
}

acq_scatter <- score_scatter(non_donors, "Acquisition", acq_col)
ret_scatter <- score_scatter(constituents, "Retention", ret_col)
upg_scatter <- score_scatter(constituents, "Upgrade", upg_col)
ltv_scatter <- score_scatter(constituents, "Lifetime Value", ltv_col)

scatter <- grid.arrange(acq_scatter, ret_scatter, upg_scatter, ltv_scatter, nrow = 2)
ggsave("../Images/DonorSearch Insights Scatter.png", scatter, width = 13, height = 9, units = "in")
scatter
```

Bar plots:
Create individual bar plots for each insight score by wealth capacity, colored by the corresponding insight rank.
```{r bar}
rank_bar <- function(df, col_name, colors) {
  rank_label <- paste(col_name, "Rank", sep=" ")
  rank_col <- gsub(" ", ".", rank_label)
  ggplot(df,aes_string(x = "Wealth.Based.Capacity", fill = rank_col)) + 
    geom_bar() + 
    scale_fill_manual(values = colors) +
    labs(x = "Wealth Based Capacity", y = "Number of Constituents", fill = rank_label) + 
    theme_bw() + 
    theme(legend.position="bottom",
          axis.text.x = element_text(size = 6, angle = 90, hjust=1, vjust = 0.5),
          panel.grid.minor = element_blank())
}

acq_bar <- rank_bar(non_donors, "Acquisition", acq_col)
ret_bar <- rank_bar(constituents, "Retention", ret_col)
upg_bar <- rank_bar(constituents, "Upgrade", upg_col)
ltv_bar <- rank_bar(constituents, "Lifetime.Value", ltv_col)

bar <- grid.arrange(acq_bar, ret_bar, upg_bar, ltv_bar, nrow = 2)
ggsave("../Images/DonorSearch Insights Bar.png", bar, width = 13, height = 9, units = "in")
bar
```

Plot Number of Constituents by Constituency Type, colored by Donor Type
```{r alumni}
alumni <- constituents %>% group_by(Constituency.Type, Donor.Type) %>% summarise(Frequency = n())
alumni_plot <- ggplot(alumni, aes(x = Constituency.Type, y = Frequency, fill = Donor.Type, label = Frequency)) +
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("#48C2CB","#F38531")) +
  labs(x = "Constituency Type", y = "Number of Constituents", fill = "Donor Type", title = "Number of Constituents by Constituency Type") + 
  theme_classic()
ggsave("../Images/DonorSearch Insights Alumni.png", alumni_plot, width = 9, height = 13, units = "in")
alumni_plot
```

Plot Number of Constituents by Graduation Year
```{r grad}
grad_year <- constituents %>% count(Pref_Year, Donor.Type) %>%  na.omit()
grad_plot <- ggplot(grad_year, aes(x = Pref_Year, y = n, color = Donor.Type)) +
  geom_line() + 
  scale_color_manual(values = c("#48C2CB","#F38531")) +
  labs(x = "Graduation Year", y = "Number of Constituents", color = "Donor Type", title = "Number of Constituents by Graduation Year") + 
  theme_bw()
ggsave("../Images/DonorSearch Insights Graduation.png", grad_plot, width = 13, height = 9, units = "in")
grad_plot
```